---
title: "OkCupid"
author: "Project 5"
date: "November 30, 2017"
output: html_document
---

# Set-Up

Install Packages:

```{r}
#install.packages("fmsb")

library(MASS)
library(fmsb)
library(ggplot2)
```

Load Data:

```{r}
raw.data.speed.dating <- read.csv("../data/speed-dating-experiment/Speed Dating Data.csv")
load("../data/profiles.RDA")
```

# Data Description

The speed dating data set is...

# Data Filter

Relevant Data:

```{r}
#For looking at matches
select.fields <- c("match", "int_corr", "samerace", "age")
speed.dating <- raw.data.speed.dating[ , select.fields]


#For looking at individuals
select.fields2 <- c("iid", "gender", "race", "imprace", "imprelig", "age",
                    "income", "goal", "date", "go_out", "career_c", "sports", 
                    "dining", "museums", "art", "gaming", "clubbing", "reading", 
                    "music", "yoga", "attr1_1", "sinc1_1", "intel1_1", "fun1_1",
                    "amb1_1", "shar1_1")

speed.dating.individual <- raw.data.speed.dating[ , select.fields2]
speed.dating.individual <- unique(speed.dating.individual)

rm(select.fields, select.fields2)
```

Calculate matching rates for individuals and merge with individuals dataset:

```{r}
matching.rates <- raw.data.speed.dating[, c("match", "iid")]
matching.rates <- aggregate(match ~ iid, data = matching.rates, mean)

speed.dating.individual <- merge(speed.dating.individual, matching.rates, by = "iid")
```

# Data Analysis

##Preliminary Analysis:

```{r}
#For every variable 1 = yes and 0 = no
#For every rate, 10 is important and 0 is not important

#Matching analysis-----------------------------------------------------------------------

#Matching percentage
round(sum(speed.dating$match == 1)/nrow(speed.dating) * 100, 2)

#Age
boxplot(speed.dating.individual$age)

#Box plot on matches by ratings of interest correlations
par(mfrow=c(1,2))
boxplot(int_corr ~ match, data = speed.dating, main = "Correlation on ratings of interests x Match")
boxplot(speed.dating$int_corr, main = "Correlation on ratings of interests")

#Race
table(speed.dating$match, speed.dating$samerace)
length(unique(speed.dating$race))
table(speed.dating$race, speed.dating$match)

#Individual analysis (taking repeated rows)-----------------------------------------------

```

### importance of religion and race/ethnic background

```{r}
#How important it is to be with someone of the same religion or race
par(mfrow=c(1,2))
boxplot(speed.dating.individual$imprace, main = "Same race/ethnic background  rating")
boxplot(speed.dating.individual$imprelig, main = "Same religion rating")

##
## Apparently, neither aspects are so important


#Matching rates
boxplot(speed.dating.individual$match, main = "Matching rates")

## 
## OBS: didn't get so well this plot. (Joaquim)
```

### What people are looking for in their partner

```{r}
#What people are looking for in his partner
par(mfrow=c(1,1))
radar <- apply(speed.dating.individual[,c("attr1_1","sinc1_1","intel1_1","fun1_1","amb1_1","shar1_1")], 2, mean, na.rm = T)
names(radar) <- c("Attractive", "Sincere", "Inteligent", "Fun", "Ambitious", "Shared Hobbies")
radar <- rbind( rep(30,6), rep(0,6), radar)
radarchart(as.data.frame(radar), title = "What people are looking for")

par(mfrow=c(1,2))
radar <- apply(speed.dating.individual[speed.dating.individual$gender == 0, c("attr1_1", "sinc1_1", "intel1_1", "fun1_1", "amb1_1", "shar1_1")], 2, mean, na.rm = T)

names(radar) <- c("Attractive", "Sincere", "Inteligent", "Fun", "Ambitious", "Shared Hobbies")
radar <- rbind( rep(30,6), rep(0,6), radar)
radarchart(as.data.frame(radar), title = "What women are looking for")

radar <- apply(speed.dating.individual[speed.dating.individual$gender == 1, c("attr1_1", "sinc1_1", "intel1_1", "fun1_1", "amb1_1", "shar1_1")], 2, mean, na.rm = T)

names(radar) <- c("Attractive", "Sincere", "Inteligent", "Fun", "Ambitious", "Shared Hobbies")
radar <- rbind( rep(30,6), rep(0,6), radar)
radarchart(as.data.frame(radar), title = "What men are looking for")

##
## As we can see, men seem to care more if women are attractive or not, when it comes to looking for a partner. Women tend to value more intelligence in their partners.
```

### Matches statistics
```{r}
plot(density(speed.dating.individual$match), main = "matches density")

# get quantiles, mean and std dev
quantile(speed.dating.individual$match, c(0.1,0.25,0.5,0.75, .9))
mean(speed.dating.individual$match)
sd(speed.dating.individual$match)
```

### Is there any pattern among people with high match rates?

```{r}
par(mfrow=c(1,1))

speed.dating.individual$match <- speed.dating.individual$match + 0.0001 

#First Model----------------------------------------------------------------------

#Fit linear model with all variable
linear.model <- lm(match ~ ., data = speed.dating.individual[,-c(1,7)])

#Interesting results
summary(linear.model)

sum(linear.model$residuals^2)

#Second Model----------------------------------------------------------------------

#Selecting variables
linear.model <- lm(match ~ imprace + age + date + go_out + art + clubbing, 
                   data = speed.dating.individual[,-c(1,7)])

#Second model results
anova(linear.model)

summary(linear.model)

sum(linear.model$residuals^2)

#Third Model----------------------------------------------------------------------

linear.model <- lm(match ~ 0 + age + go_out + art + clubbing, 
                   data = speed.dating.individual[,-c(1,7)])

#Second model results
anova(linear.model)
summary(linear.model)
sum(linear.model$residuals^2)

linear.stdres = rstandard(linear.model)
qqnorm(linear.stdres, 
     ylab="Standardized Residuals", 
     xlab="Normal Scores", 
     main="Old Faithful Eruptions") 
qqline(linear.stdres)


#Interesting: the date variable was important on the second model, but not on the third... are they highly correlated? or are people lying?

calculate.correlation <- speed.dating.individual[ , c("date", "go_out")]
calculate.correlation <- na.omit(calculate.correlation)

cor(calculate.correlation)
```

### Can we predict whether 2 people will match? 

```{r}
lda.model <- lda(match ~ ., data = speed.dating, CV = T)

table(lda.model$class)
```


### income and matches
Is there any correlation between a person's income and his percentage of matches?
```{r}

plot( x = speed.dating.individual$match, y = speed.dating.individual$income )

cor( x = speed.dating.individual$match, y = as.numeric(speed.dating.individual$income) )

```
Apparently, it doesn't appear to exist a relation between these two factors.


### who usually goes out more on dates? men or women?
```{r}
df_n_dates <- speed.dating.individual[,c("gender","date")]
df_n_dates[ which( df_n_dates$gender == 0 ), ]$gender <- "female"
df_n_dates[ which( df_n_dates$gender == 1 ), ]$gender <- "male"

# plot histograms
dat <- data.frame(xx = c(runif(100,20,50),runif(100,40,80),runif(100,0,30)),yy = rep(letters[1:3],each = 100))

ggplot(df_n_dates,aes(x = date)) + 
    geom_density(data=subset(df_n_dates,gender == 'female'),fill = "red", alpha = 0.2) +
    geom_density(data=subset(df_n_dates,gender == 'male'),fill = "blue", alpha = 0.2)

mean( df_n_dates[ which( df_n_dates$gender == 'female' & !is.na(df_n_dates$date) ), ]$date)
mean( df_n_dates[ which( df_n_dates$gender == 'male' & !is.na(df_n_dates$date) ), ]$date)
```
As we can see, apparently women go more often on dates than men.


Is there any pattern among people with high match rates?

Can we predict someone's matching rate?

Can we cluster matches?

Can we predict whether 2 people will match? 

